---
title: "Mountain Yellow-Legged Frogs in the Sierra Nevada"
description: |
  Visualization of Mountain yellow-legged frog populations in the Sierra Nevada.
author:
  - name: Roni Shen
date: 02-23-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load packages
library(tidyverse)
library(lubridate)
library(here)
library(patchwork)
library(readxl)

# read in dataset
amphibians <- read_excel(here("_posts","2021-02-23-frogs","sierra_amphibians.xlsx"))
```

### Part 1: Creating a graph of mountain yellow-legged frog count per year by life stage

```{r life stage plot, echo=TRUE, message=FALSE, warning=FALSE}

# wrangle data and create a subset
amphibians_ramu <- amphibians %>% 
  filter(amphibian_species == "RAMU", 
         amphibian_life_stage %in% c("Adult", "SubAdult", "Tadpole")) %>% 
  mutate(date = ymd(survey_date),
         survey_year = year(survey_date),
         amphibian_life_stage = factor(amphibian_life_stage, 
                                       levels = c("Adult", 
                                                  "SubAdult", 
                                                  "Tadpole"))) 

# create a plot with customized aesthetics
plot_a <- ggplot(data = amphibians_ramu, aes(x = survey_year, y = amphibian_number, 
                                   fill = amphibian_life_stage)) +
  geom_col(position = "stack") +
  theme_classic() +
  scale_fill_manual(values = c("Adult" = "deepskyblue3", 
                                "SubAdult" = "salmon1", 
                                "Tadpole" = "darkorchid")) +
  scale_y_continuous(expand = c(0,0)) +
  labs(subtitle = "Count per year by life stage", 
       x = "Year", 
       y = "Count", 
       fill = "Life stage") +
  theme(legend.position = c(.4,.7))
```

### Part 2: Creating a graph of mountain yellow-legged frog by total counts of adults and subadults observed in the 5 lakes with highest totals

```{r lake plot, echo=TRUE, message=FALSE, warning=FALSE}

# wrangle data from previous subset and create a new subset
amphibians_lakes <- amphibians_ramu %>% 
  filter(amphibian_life_stage != "Tadpole") %>% 
  group_by(lake_id) %>% 
  summarise(total = sum(amphibian_number)) %>% 
  ungroup() %>% 
  slice_max(total, n = 5) %>% 
  mutate(lake_id = fct_inorder(as.character(lake_id)))

# create a plot with customized aesthetics  
plot_b <- ggplot(data = amphibians_lakes, aes(x = lake_id, y = total, fill = lake_id)) +
  geom_col(show.legend = FALSE) +
  theme_classic() +
  scale_y_continuous(expand = c(0,0)) +
  labs(subtitle = "Total count by top 5 lakes with highest observed counts", 
       x = "Lake ID number",
       y = "Total count (adults and subadults)")
```

### Part 3: Combining plots for mountain yellow-legged frogs with yearly totals for 3 life stages (part 1) and total counts for adults and subadults in the 5 lakes with highest observed counts (part 2)

```{r combined plot, echo=TRUE, message=FALSE, warning=FALSE, fig.width=10}

# combine 2 plots with {patchwork} and customized aesthetics
combined_plot <- (plot_a | plot_b) +
  plot_layout(widths = c(1,1)) +
  plot_annotation(title = "Mountain yellow-legged frog populations in the Sierra Nevada",
                  caption = "Figure 1. Mountain yellow-legged frogs totals across 3 life stages (left) and total counts for adults and subadults in the 5 lakes with highest observed counts (right).") &
  theme(plot.caption = element_text(hjust = 0, size = 10),
        plot.title = element_text(hjust = .25, size = 18))

# view combined plot
combined_plot
```

