---
title: "Binary Logistic Regression: Palmetto spp"
description: |
  A binary logistic regression comparing two Florida palmetto species.
author:
  - name: Roni Shen
date: 02-23-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
library(patchwork)
library(broom)
library(GGally)
library(jtools)

palmetto <- read_csv(here("_posts","2021-02-23-palmettoblr", "palmetto.csv"), 
    col_types = cols(
      species = col_factor(),
      treatment = col_double(),
      new_lvs = col_double(),
      canopy = col_double(),
      lf_long = col_double(),
      comments = col_character()
    )
  ) %>% 
  select(year, species, site, height, length, width, green_lvs)
  
levels(palmetto$species) <- c("S. repens", "S. etonia")

```

## Data Visualizations and Binary Logistic Regression Model of Two Florida Palmetto Species



### Distributions

```{r fig.cap="Figure 1. Distributions of palmetto height, canopy length, canopy width, and number of green leaves by species."}

height <- ggplot(palmetto, aes(x = height)) +
  geom_freqpoly(aes(color = species), binwidth = 10) +
  labs(y = "Height") +
  theme(legend.position = "none",
        axis.title.x = element_blank())

length <- ggplot(palmetto, aes(x = length)) +
  geom_freqpoly(aes(color = species), binwidth = 10) +
  labs(y = "Canopy Length") +
  theme(legend.position = "none",
        axis.title.x = element_blank())

width <- ggplot(palmetto, aes(x = width)) +
  geom_freqpoly(aes(color = species), binwidth = 10) +
  labs(y = "Canopy Width") +
  theme(legend.position = "none",
        axis.title.x = element_blank())

leaves <- ggplot(palmetto, aes(x = green_lvs)) +
  geom_freqpoly(aes(color = species), binwidth = 1) +
  labs(y = "Number of Green Leaves") +
  theme(legend.position = "none",
        axis.title.x = element_blank())

(width | length) / (height | leaves) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = 'right') & 
  labs(color = "Species")
```
From figure 1, we can see the relative frequency distributions and variance of each of the four variables of interest between two palmetto species and their relative similarities.


### Pairs Plots

```{r fig.cap="Figure 2. Pairs plots for multivariate correlations and distributions between palmetto variables of the two species."}
palmetto %>% 
  select(species, height:green_lvs) %>% 
  ggpairs(aes(color = species)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```
From figure 2, we can see that the two palmetto species are very similar and have significant correlations across multiple variables of measure. 



### Binary logistic regression

```{r}

palmetto_blr <- glm(species ~ height + width + length + green_lvs, 
                    data = palmetto,
                    family = "binomial")

# summary table
palmetto_tidy <- tidy(palmetto_blr) %>% 
  select(!"statistic")  

knitr::kable(palmetto_tidy, col.names = c("Variable", "Coefficient", "Standard Error",
                                          "p Value"))

```


### Model 

Table 1. Incidents of correct and incorrect predictions of palmetto species based on binary logistic model of predictor variables.


```{r, results='asis'}
# Model fitting
blr_fitted <- palmetto_blr %>% 
  augment(type.predict = "response") %>% 
  mutate(etonia = ifelse(.fitted >= .5, 1, 0),
         repens = ifelse(.fitted < .5, 1, 0)) %>%
  group_by(species) %>% 
  summarise(etonia = sum(etonia, na.rm = T),
            repens = sum(repens, na.rm = T),
            `% correct` = ifelse(species == "S. repens", repens / n() * 100, 
                                 etonia / n() * 100)) %>% 
  distinct()

knitr::kable(blr_fitted, col.names = c("Species", "Predicted as S. etonia", 
                                       "Predicted as S. repens", "% Correct"))

```




