---
title: "Text and Sentiment Analysis of Lyrics from Abbey Road (The Beatles)"
description: |
  Examples of text and sentiment analysis using R.
author:
  - name: Roni Shen
date: 02-23-2021
output:
  distill::distill_article:
    self_contained: false
    code_folding: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidytext)
library(textdata)
library(ggwordcloud)
library(pdftools)
library(here)

```

```{r}
beatles <- pdf_text(here("_posts","2021-02-23-textanalysis","beatles.pdf"))

beatles_tidy <- data.frame(beatles) %>% 
  mutate(lyrics = str_split(beatles, pattern = '\\n')) %>% 
  unnest(lyrics) %>% 
  mutate(lyrics = str_trim(lyrics))
```

```{r message=FALSE, warning=FALSE}
beatles_df <- beatles_tidy %>% 
  mutate(song = case_when(str_detect(lyrics, "Song") ~ lyrics, 
                          TRUE ~ NA_character_)) %>% 
  fill(song) %>% 
  separate(col = song, into = c(" ", "song"), sep = " ") %>% 
  mutate(song = as.numeric(song))

beatles_tokens <- beatles_df %>% 
  unnest_tokens(word, lyrics) %>% 
  select(-beatles) %>% 
  anti_join(stop_words) %>% 
  filter(!str_detect(word, "song")) %>% 
  mutate(word = str_replace(word, "[0-9-]+", NA_character_))
```

# Word Cloud 
## Top 5 Words by Song

```{r message=FALSE, warning=FALSE}
word_count <- beatles_tokens %>%  
  count(song, word) %>% 
  group_by(song) %>% 
  arrange(-n) %>% 
  slice(1:5)

ggplot(data = word_count, aes(label = word)) +
  geom_text_wordcloud(aes(color = n, size = n), shape = "circle") +
  scale_size_area(max_size = 15) +
  scale_color_gradientn(colors = c("green","purple","red")) +
  theme_minimal()
```


# Sentiment Analysis 
## By Song

#### Track List:
Track 1: Come Together

Track 2: Maxwell's Silver Hammer

Track 3: Something

Track 4: Oh! Darling

Track 5: Octopus's Garden

Track 6: I Want You (She's so Heavy)

Track 7: Here Comes the Sun

Track 8: Because

Track 9: You Never Give Me Your Money

Track 10: Sun King

Track 11: Mean Mr. Mustard

Track 12: Polythene Pam

Track 13: She Came In Through the Bathroom Window

Track 14: Golden Slumbers

Track 15: Carry That Weight

Track 16: The End

Track 17: Her Majesty


```{r message=FALSE, warning=FALSE, fig.width=10, fig.height=7}
beatles_nrc <- beatles_tokens %>% 
  inner_join(get_sentiments("nrc")) %>%  
  count(song, sentiment)

ggplot(data = beatles_nrc, aes(x = sentiment, y = n, fill = sentiment)) +
  geom_col() +
  facet_wrap(~song) +
  coord_flip()
  
```

Source: The Beatles (1969). *Abbey Road*. Apple Records.